Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-12 writedown/period.noneq.msm/paper.tex
revise by CS

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing}.
It allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $k$ core sets $C_1,\ldots, C_k\subset S$, and we denote $C=S\setminus\cup_j C_j$. Following
Ref.~\cite{djurdjevac2010markov}, Thm. 3.1, the coarse grained transition
matrix $\hat{P}$ then has to be computed as follows:
\begin{enumerate}
\item For the process associated with $P$ compute the forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$. This can be done by solving the linear equations
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\label{qfwd}\\
q^\fwd_j(i) & =  1,\quad i\in C_j\nonumber\\
q^\fwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\label{qbwd}\\
q^\bwd_j(i) & =  1,\quad i\in C_j\nonumber \\
q^\bwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu_i P_{ij}/\mu_j$.
\item Compute $\hat{\mu}(j)=\sum_{i\in S} q^\bwd_j(i)\mu(i)$,
  which is the invariant measure of the MSM
\item Construct the MSM transition matrix $\hat{P}$ according to
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u_i v_i \mu_i$.
\end{enumerate}

\noindent\textbf{Remark 1:}
Alternatively, the forward and
backward committors can be \emph{sampled} by means of the following procedure: The forward
committor $q^\fwd_j(i)$ is defined as the probability of visiting coreset
$C_j$ next conditioned on being at state $i$.  The
backward committor $q^\bwd_j(i)$  is defined as the probability
of last coming from $C_j$ conditioned on being at state $i$.

% \subsection{Building the Markov State Model from the NEMD simulation}
\noindent\textbf{Remark 2:}
Sometimes, it is not easy to have an explicit expression for the committors, however, we
can still build the MSM by using the following procedure: Firstly compute the
matrixes $\hat T$ and $\hat M$ by
\begin{align}\label{eq:msm-ht}
  \hat M_{jk} &= P [\,h_{mT}(C_k) < h_{mT}(\cup_{l\neq k}C_l) \,\vert\, \hat x_{mT} = j\,],\\\label{eq:msm-hm}
  \hat T_{jk} &= P [\,h_{mT+T}(C_k) < h_{mT+T}(\cup_{l\neq k}C_l) \,\vert\, \hat x_{mT} = j\,],
\end{align}
where $\hat x_{mT}$ is milestone process corresponding to the NEMD trajectory $x_{mT}$ (notice here
only integral periods $mT$ along the NEMD trajectories are used for building the milestone process).
In Eq.~\eqref{eq:msm-ht} and \eqref{eq:msm-hm}, the time parameterized first hitting time of  set $A$ is defined by:
\begin{align}
  h_{t}(A) = \inf_{s\geq t}\{ s : x_s \in A \}.
\end{align}
It can be shown that (see Appendix~\ref{sec:app-prove} for details):
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}.
\end{align}
% \recheck{We may need to write why do these formula works, because in literature it is only available for reversible processes.}
Therefore we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{ij} - \hat M_{ij}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \hat P_{jk}, \quad j\neq k
\end{align}
By Eq.~\eqref{eq:msm-ht} and \eqref{eq:msm-hm}, we have $\sum_i\hat M_{ij} = \sum_i\hat T_{ij} = 1$, then
$\sum_k(\hat M_{jk}- \hat T_{jk}) = 0$. By the second line of \eqref{eq:msm-tmatrix-00}, $\sum_k\hat P_{jk} = 1$, so using \eqref{eq:msm-tmp26} we have
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$.
In the matrix expression the MSM transition matrix $\hat P$ is given by
\begin{align}\label{eq:msm-tmatrix-10}
  \hat P = \hat T - \hat M + \id
\end{align}
% Given coresets $\{C_1, \cdots, C_k\}$, the procedure \eqref{eq:msm-ht}--\eqref{eq:msm-tmatrix-10}
% If assuming that we have the coresets $\{C_1, \cdots, C_k\}$, no matter how they are identified,
In constrast to Eq.~\eqref{eq:msm-tmatrix-00}, which needs the intermediate spacial and
temporal discretization (i.e.~the Floquet transition matrix $P$), 
Eq.~\eqref{eq:msm-ht}, \eqref{eq:msm-hm} and \eqref{eq:msm-tmatrix-10} allows a direct derivation of the MSM from the original NEMD
trajectories, which substantially simplified the MSM building.
However, in the present paper, we present both the approachs,
because the former contains more information regarding the original dynamics in $P$ and $L(t)$ than in $\hat P$.
Moreover, the coreset identification may need a good spacial discretization to the original NEMD trajectories, which is
usually consisten with that used to go from $\ml(T)$ to $L(t)$. In the end, the MSM
built from both approachs should be consistent, taking no account of the statistical uncertainty in the samplings.

@@

Following the process described in Sec.~\ref{sec:build-msm}, we are able to build a 
three state MSM for the externally driven alanine dipeptide system,
where the quasi-stationary probability distribution $\mu$, the three core sets, and the forward and backward committors are
estimated as described above, and the MSM transition matrix $\hat P$ is then evaluated using Eq.~\eqref{eq:msm-tmatrix-00}.
Alternatively, the MSM  transition matrix $\hat P$ is calculated directly from the NEMD trajectories by Eq.~\eqref{eq:msm-tmatrix-10}.
The leading eigenvalues of $P=\myphi^T(T)$ are compared with those of $\hat P$ in
Tab.~\ref{tab:tmp1}.
Without surprising, the two approaches for MSM building are consitent.
The MSM is able to accurately reproduce
the largest non-trivial eigenvalue, which means a precise reproduction
of the longest non-trivial implied time-scale. The accuracy of the second non-trivial
time-scale is not as good as the first, but is still acceptable. The reason
for the lower accuracy is that
the corresponding time scale is 26.5~ps (calculated by $-T/\log(\lambda_2)$),
which is NOT significantly longer than the temporal resolution given by the period $T=10$~ps of the external driving force.
It is worth noting that although the discretized process $\tilde X_m$ is irreversible,
the MSM built out of it is almost reversible:
The magnitude of the anti-symmetric part of the
matrix $\textrm{diag}(\hat \mu)\cdot \hat P$ is only of order $10^{-4}$.

\begin{table}
  \centering
  \caption{
    Comparison of second and third eigenvalues of $\vect P$ and $3\times 3$ MSM transition matrix $\hat{P}$ from two approaches.
  }
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}cc rrr}\hline\hline
      & method &  $\lambda_2$ & $\lambda_3$ & $\lambda_4$ \\\hline
    $P$         & --         &0.907  &0.686 & 0.553       \\
    $\hat P$    &\eqref{eq:msm-tmatrix-00} & 0.911  &0.724 & --       \\
    $\hat P$    &\eqref{eq:msm-tmatrix-10} & 0.909  &0.712 & --       \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp1}
\end{table}

@@

In fact, $\hat{P}$ can be considered as the fingerprint of the long-term kinetics (cf. \cite{A19-39,PrinzKellerNoe_PCCP11_Perspective}) of alanine dipeptide in an oscillatory electric field.
In order to provide further validation of this statement, we study  time-dependent expectation values of
the form
\begin{align}
  \mathcal A(t) = \langle A(i)\rangle_t = \sum_{i\in S} A(i) p(i,t),
\end{align}
where $p_i(t)=\prob(X_t=i)$ denotes the probability to be in set $\Omega_i$ at time $t$ as governed by the Master equation, and the observable $\mathcal A$ is spanned by the backward
committors, i.e.,
\begin{align}\label{A}
  A(i) = \sum_{l=1}^k \alpha_l q^\bwd_l(i).
\end{align}
Then
\begin{align}\nonumber
  \mathcal A(t) &=
  \sum_{i\in S} \sum_{l=1}^k \alpha_l q^\bwd_l(i)  p(i,t)=\sum_{i\in S} \sum_{l=1}^k \alpha_l \prob (\hat X_t = l \vert X_t = i) \prob (X_t = i) \\
  & =
  \sum_{i\in S} \sum_{l=1}^k \alpha_l \prob (\hat X_t = l ,X_t = i)  =
  \sum_{l=1}^k \alpha_l \prob (\hat X_t = l) 
 =
   \sum_{l=1}^k \alpha_l \,\hat p (l, t), \label{eq:num-28}
\end{align}
where the time-dependent probability $\hat{p}(i,t)$ of being assigned to MSM macrostate $i$ at time $t$ can be computed  by means of the MSM via simple matrix multiplications:
\begin{align}\label{eq:num-29}
  \hat p(i, t+T) = \sum_{j\in S} \hat p(j,t)\hat P_{ji}.
\end{align}
In Fig.~\ref{fig:num-7} we compare the numerical calculation of $\hat p (l, mT), \ m\in\mathbb N$ from NEMD and MSM calculations.
In the NEMD case, the backward committor and the probability density $\hat p (l, mT) = \sum_{i\in S}  q^\bwd_l(i)  p(i,mT) $
are estimated directly from the
NEMD trajectories. For the MSM, the projection of the initial probability is applied, $\hat p (l, 0) = \sum_{i\in S}  q^\bwd_l(i)  p(i,0) $, then
the time-dependent probability at $mT$ is generated by Eq.~\eqref{eq:num-29}, i.e., by simple matrix multiplication. The agreement is almost perfect.
% The projected probability, e.g.~$\hat p (l', mT)$ is calculated
% by Eq.~\eqref{eq:num-28} letting $\alpha_l = \delta_{l'l}$.

@@

\section{Proves for Equations~\eqref{eq:msm-ht-1} and \eqref{eq:msm-hm-1} }
\label{sec:app-prove}

We firstly define $\hat q^\bwd_j (i) = P (X_t = i \vert \hat X_t =
j)$, and always assume $t = mT$. Then due to the Bayes' Theorem we have
\begin{align}
  \hat q^\bwd_j (i) =
  \frac{P (\hat X_t = j \vert X_t = i) P(X_t = i)}{ P(\hat X_t = j) } =
  \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(j)}
\end{align}
Then
\begin{align*}
  P& [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, \hat X_t = j\,]\\
  &=
  \sum_{i=1}^n
  \frac{P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) , X_t = i, \hat X_t = j\,]}
  { P(\hat X_t = j )} \\
  &=
  \sum_{i=1}^n
  {P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, X_t = i, \hat X_t = j\,]}\,
  \hat q^\bwd_j(i) \\
  &=
  \sum_{i=1}^n
  {P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, X_t = i\,]}\,
  \hat q^\bwd_j(i)  \\
  &=
  \sum_{i=1}^n
  q^\fwd_k(i) \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(i)} \\
  &=
  \frac{\langle q_j^\bwd, q_k^\fwd\rangle_\mu}{\hat \mu(j)}
\end{align*}
The third equation holds because of the Markovianity of the process $X_t$.
The fourth equation is due to the definition of the forward committor.
This proves Eq.~\eqref{eq:msm-ht-1}.
To prove Eq.~\eqref{eq:msm-hm-1}, we firstly need
\begin{align*}
  P & [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \,\vert\, X_t = i\,]\\
  &=
  \sum_{l=1}^n
  \frac{P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l), X_{t+T} = l, X_t = i\,]}
  {X_t = i} \\
  &=
  \sum_{l=1}^n
  {P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \vert X_{t+T} = l, X_t = i\,]}
  {P( X_{t+T} = l \vert X_t = i)}\\
  & =
  \sum_{l=1}^n
  {P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \vert X_{t+T} = l\,]}
  {P( X_{t+T} = l \vert X_t = i)} \\
  &=
  \sum_{l=1}^n
  q^\fwd_k(l) P_{il}
\end{align*}
We used the Markovianity of $X_T$ at the third equation,
and the time-homogeneity at the fourth equation.
Therefore, following the same procedure proving Eq.~\eqref{eq:msm-ht-1}, we have
\begin{align*}
  P& [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \,\vert\, \hat X_t = j\,]\\
  &=
  \sum_{i=1}^n\sum_{l=1}^n
  q^\fwd_k(l) P_{il}
  \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(i)} \\
  &=
  \frac{\langle q_j^\bwd, P q_k^\fwd\rangle_\mu}{\hat \mu(j)}  
\end{align*}
This proves Eq.~\eqref{eq:msm-hm-1}.

>>>>

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $k$ core sets $C_1,\ldots, C_k\subset S$ that are appropriate metastable sets. 
Following \cite{A19-31,schuette2011markov} the process $(X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $k$ states associated with the sets $C_j$, $j=1,\ldots,k$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(X_m)$, i.e., trajectories of $(X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\[
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j|X_m=i),
\] 
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
From the last equation we get that the stationary distribution of the milestone process is given by
\[
\hat{\mu}(j)=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\]
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}(j)$.

Following \cite{discreteTPT} one also has to consider the forward committor $q^\fwd_j(i)$ identical to the probability  that conditioned on $X_m=i$ the next core set to be hit will be $C_j$.
The forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$ can be computed from $P$ by solving the linear equations \cite{discreteTPT}
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\label{qfwd}\\
q^\fwd_j(i) & =  1,\quad i\in C_j\nonumber\\
q^\fwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $C=S\setminus\cup_j C_j$, and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\label{qbwd}\\
q^\bwd_j(i) & =  1,\quad i\in C_j\nonumber \\
q^\bwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu_i P_{ij}/\mu_j$.

Following
Ref.~\cite{djurdjevac2010markov}, Thm. 3.1, the one-step transition
matrix $\hat{\vect P}$ associated with the milestone process $\hat{X}_m)$ is given by
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jl}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_l \rangle_\mu,\qquad j\not= l, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{l\not=j} \hat{\vect P}_{jl}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u_i v_i \mu_i$. More precisely, we have
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jl}=\prob_\mu\Big(\hat{X}_{m+1}=l\vert \hat{X}_m=j\Big).
\end{equation}
In general the milestone process need not be a Markov process. The results in \cite{A1-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}(l,0)$ of being in state $l$ at time $t=0$ then its evolution $\hat{p}(l,t)$ in time is discrete in multiples of period $T$ and given by
\recheck{
\begin{equation}\label{eq:num-29}
\hat{p}(l,T)=\sum_j \hat{p}(j,0)\,\,\hat{\vect P}_{jl}.
\end{equation}}
In our approach $p(j,mT)$ is a good approximation of $\prob(\hat{X}_m=j)$.

\textbf{Remark 1:} Our definition of the milestoning process in terms of the process $(X_m)$ generated by $P$ guarantees that we can directly compute $\hat{P}$ via (\ref{hatP_prob})  from trajectories of $(X_m)$ without computing the committor functions. 
This is of importance if the spatial discretization underlying $(X_m)$ is fine enough, because then the kinetics of $(X_m)$
approximates the original kinetics of $(x_{mT})$ so that we can directly compute   $\hat{P}$ via (\ref{hatP_prob})  from NEMD trajectories without computing $P$ first (which substantially simplified the MSM building if the core sets are already known).

\textbf{Remark 2:}
Following \cite{A19-31} we can also define another pair of stochastic MSM matrices: 
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)},
\end{align}
that are connected to $\hat{P}$ by the following identity
\[
\hat{P}=\hat{T}-\hat{M}+\id,
\]
That can be seen by means of direct computation: For the off-diagonal entries we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{ij} - \hat M_{ij}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \hat P_{jk}, \quad j\neq k
\end{align}
Stochasticity yields
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$ for the diagonal entries.  Furthermore, as shown in the Appendix $\hat{T}$ as well as $\hat{M}$ can be computed from trajectories without need to have the committors. 
The importance of the pair $\hat{T}$ and $\hat{M}$ for MSM building comes from the following observation: The main NEMD relaxation timescales are given by the dominant eigenvalues of $P$ \cite{A19-31, A19-1}.
These dominant eigenvalues can be approximated by discretizing the related eigenvalue problem $Pu=\lambda u$ by means of a Galerkin approximation with the finite dimensional ansatz space 
spanned by the forward committors $q^+_j$, $j=1,\ldots,k$ together  with the finite dimensional test function space spanned by the backward committors $q^-_j$, $j=1,\ldots,k$ (test functions multiplied from the left by the inner product $\langle\cdot,\cdot\rangle_\mu$). The thus discretize eigenproblem takes the form  of a generalized eigenproblem
\[
\hat{T}\hat{u}=\hat{\lambda}\hat{M}\hat{u}.
\]
For the reversible case it is known that its $k$ eigenvalues $\hat{\lambda}$ are very good approximation of the dominant eigenvalues $\lambda$ of the original problem if the core sets are proper metastable sets \cite{Eigenvalues}. Whether this is true for the non-reversible case is not known yet, but if the deviation from reversibility is weak and the dominant eigenvalues of $P$ are real-valued then the results should hold analogously, see \cite{A19-31}, Thm. 4.19.

The pair $\hat{T}$ and $\hat{M}$, or the matrix $\hat{M}^{-1}\hat{T}$ (if all of its entries are positive such that it is a stochastic matrix), thus can also be taken as MSM transition matrices. In the case of  $\hat{T}$ and $\hat{M}$ the MSM modelling error results from Galerkin discretization, while the MSM modelling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$. The lesson learnt in the reversible case are that $\hat{P}$ is better suited to approximate the molecular kinetics on large timescales while $\hat{T}$ and $\hat{M}$ should be used for approximating eigenvalues and long relaxation timescales.

% Given coresets $\{C_1, \cdots, C_k\}$, the procedure \eqref{eq:msm-ht}--\eqref{eq:msm-tmatrix-10}
% If assuming that we have the coresets $\{C_1, \cdots, C_k\}$, no matter how they are identified,

@@

Following the process described in Sec.~\ref{sec:build-msm}, we are able to build a 
three state MSM for the externally driven alanine dipeptide system,
where the quasi-stationary probability distribution $\mu$, the three core sets, and the forward and backward committors are
estimated as described above, and the MSM transition matrices $\hat P$, $\hat{M}$, and $\hat{T}$ is then evaluated using Eq.~\eqref{eq:msm-tmatrix-00}and \eqref{eq:msm-ht-1}.
Alternatively, the MSM  transition matrix $\hat P$ is calculated directly from the NEMD trajectories using Eq.~\eqref{hatP_prob} (very good agreement with the one computed from the committors).
The leading eigenvalues of $P=\myphi^T(T)$ are compared with those of $\hat P$ and $\hat{M}^{-1}\hat{T}$ in
Tab.~\ref{tab:tmp1}.
Without surprising, the two approaches for MSM building are consitent.
The MSM is able to accurately reproduce
the largest non-trivial eigenvalue, which means a precise reproduction
of the longest non-trivial implied time-scale. The accuracy of the second non-trivial
time-scale is not as good as the first, but is still acceptable (even good for $\hat{M}^{-1}\hat{T}$). The reason
for the lower accuracy is that
the corresponding time scale is 26.5~ps (calculated by $-T/\log(\lambda_2)$),
which is NOT significantly longer than the temporal resolution given by the period $T=10$~ps of the external driving force.
Numerically the two MSM transition matrices are:
\[
\hat{M}^{-1}\hat{T}=\left(\begin{array}{ccc}
   0.866 &   0.128 &   0.006 \\
    0.183  &  0.783  &  0.034\\
    0.018   & 0.067 &   0.915
\end{array}
\right),\qquad \hat{P}=\left(\begin{array}{ccc}
   0.888 &   0.105 &   0.007 \\
    0.149  &  0.822  &  0.029\\
    0.021   & 0.055 &   0.923
\end{array}
\right),
\]
from which we see that the longest relaxation timescale results from  the dominant metastability of the left-handed alpha-helix conformations of the peptide.

It is worth noting that although the discretized process $\tilde X_m$ is irreversible,
the MSM built out of it is almost reversible:
The magnitude of the anti-symmetric part of the
matrix $\textrm{diag}(\hat \mu)\cdot \hat P$ is only of order $10^{-4}$.

\begin{table}
  \centering
  \caption{
    Comparison of second and third eigenvalues of $\vect P$ and $3\times 3$ MSM transition matrices $\hat{P}$ and $\hat{M}^{-1}\hat{T}$, respectively, from the two different MSM approaches.
  }
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}c rrr}\hline\hline
      &  $\lambda_2$ & $\lambda_3$ & $\lambda_4$ \\\hline
    $P$                 &0.907  &0.686 & 0.553       \\
    $\hat P$     & 0.911  &0.724 & --       \\
    $\hat{M}^{-1}\hat{T}$    & 0.903  &0.662 & --       \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp1}
\end{table}

@@

In fact, $\hat{P}$ can be considered as the fingerprint of the long-term kinetics (cf. \cite{A19-39,PrinzKellerNoe_PCCP11_Perspective}) of alanine dipeptide in an oscillatory electric field.
In order to provide further validation of this statement, we study  time-dependent expectation values of
the form
\begin{align}
  \mathcal A(t) = \langle A(i)\rangle_t = \sum_{i\in S} A(i) p(i,t),
\end{align}
where $p_i(t)=\prob(X_t=i)$ denotes the probability to be in set $\Omega_i$ at time $t$ as governed by the Master equation, and the observable $\mathcal A$ is spanned by the backward
committors, i.e.,
\begin{align}\label{A}
  A(i) = \sum_{l=1}^k \alpha_l q^\bwd_l(i).
\end{align}
Then
\begin{align}\nonumber
  \mathcal A(t) &=
  \sum_{i\in S} \sum_{l=1}^k \alpha_l q^\bwd_l(i)  p(i,t)=\sum_{i\in S} \sum_{l=1}^k \alpha_l \prob (\hat X_t = l \vert X_t = i) \prob (X_t = i) \\
  & =
  \sum_{i\in S} \sum_{l=1}^k \alpha_l \prob (\hat X_t = l ,X_t = i)  =
  \sum_{l=1}^k \alpha_l \prob (\hat X_t = l) 
 =
   \sum_{l=1}^k \alpha_l \,\hat p (l, t), \label{eq:num-28}
\end{align}
where the time-dependent probability $\hat{p}(i,t)$ of being assigned to MSM macrostate $i$ at time $t$ can be computed  by means of the MSM via simple matrix multiplications using (\ref{eq:num-29}).

In Fig.~\ref{fig:num-7} we compare the numerical calculation of $\hat p (l, mT), \ m\in\mathbb N$ from NEMD and MSM calculations.
In the NEMD case, the backward committor and the probability density $\hat p (l, mT) = \sum_{i\in S}  q^\bwd_l(i)  p(i,mT) $
are estimated directly from the
NEMD trajectories. For the MSM, the projection of the initial probability is applied, $\hat p (l, 0) = \sum_{i\in S}  q^\bwd_l(i)  p(i,0) $, then
the time-dependent probability at $mT$ is generated by Eq.~\eqref{eq:num-29}, i.e., by simple matrix multiplication. The agreement is almost perfect.
% The projected probability, e.g.~$\hat p (l', mT)$ is calculated
% by Eq.~\eqref{eq:num-28} letting $\alpha_l = \delta_{l'l}$.

@@

\section{Computation of $\hat{T}$ and $\hat{M}$ directly from trajectories}
\label{sec:app-prove}

In order to show how to compute $\hat{T}$ and $\hat{M}$ directly from trajectories let us start by denoting the first hitting time of set $A$ starting at time $t$ by $h_t(A)$.
In addition we define $\hat q^\bwd_j (i) = P (X_t = i \vert \hat X_t =
j)$, and always assume $t = mT$. Then due to the Bayes' Theorem we have
\begin{align}
  \hat q^\bwd_j (i) =
  \frac{P (\hat X_t = j \vert X_t = i) P(X_t = i)}{ P(\hat X_t = j) } =
  \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(j)}
\end{align}
Then
\begin{align*}
  P& [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, \hat X_t = j\,]\\
  &=
  \sum_{i=1}^n
  \frac{P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) , X_t = i, \hat X_t = j\,]}
  { P(\hat X_t = j )} \\
  &=
  \sum_{i=1}^n
  {P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, X_t = i, \hat X_t = j\,]}\,
  \hat q^\bwd_j(i) \\
  &=
  \sum_{i=1}^n
  {P [\,h_t(C_k) < h_t(\cup_{l\neq k} C_l) \,\vert\, X_t = i\,]}\,
  \hat q^\bwd_j(i)  \\
  &=
  \sum_{i=1}^n
  q^\fwd_k(i) \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(i)} =\frac{\langle q_j^\bwd, q_k^\fwd\rangle_\mu}{\hat \mu(j)}={\hat M}_{jk}
\end{align*}
where the third equation holds because of the Markovianity of the process $X_t$, and the fourth equation is due to the definition of the forward committor.
The interpretation of the result is that when starting with the milestone process in state $j$ we have to determine the fraction of all trajectories that hit core set $C_k$ first of all core sets to estimate $\hat{M}_{jk}$.

To see the same for $\hat{T}$, we first need
\begin{align*}
  P & [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \,\vert\, X_t = i\,]\\
  &=
  \sum_{l=1}^n
  \frac{P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l), X_{t+T} = l, X_t = i\,]}
  {X_t = i} \\
  &=
  \sum_{l=1}^n
  {P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \vert X_{t+T} = l, X_t = i\,]}
  {P( X_{t+T} = l \vert X_t = i)}\\
  & =
  \sum_{l=1}^n
  {P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \vert X_{t+T} = l\,]}
  {P( X_{t+T} = l \vert X_t = i)}  = \sum_{l=1}^n
  q^\fwd_k(l) P_{il}
\end{align*}
We used the Markovianity of $X_T$ at the third equation,
and the time-homogeneity in the fourth equation.
Therefore, following the same procedure as before we have
\begin{align*}
  P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \,\vert\, \hat X_t = j\,]=
  \sum_{i=1}^n\sum_{l=1}^n
  q^\fwd_k(l) P_{il}
  \frac{q^\bwd_j(i) \mu(i)}{\hat \mu(i)} = \frac{\langle q_j^\bwd, P q_k^\fwd\rangle_\mu}{\hat \mu(j)}  = {\hat T}_{jk}
\end{align*}
Thus, when starting with the milestone process in state $j$ at some time $t$ we have to determine the fraction of all trajectories of length at least $T$ that hit core set $C_k$ first of all core sets to estimate $\hat{T}_{jk}$.

\section{The equivalence}
\begin{align*}
  \hat T_{jk} &=
  P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l) \,\vert\, \hat X_t = j\,]\\
  &=
  \frac{P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l), \hat X_t = j\,] }{P(\hat X_t = j)}\\
  & =
  \sum_l \frac{P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l), \hat X_{t+T} = l, \hat X_t = j\,] }{P(\hat X_t = j)}\\
  & =
  \sum_l P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l)\vert \hat X_{t+T} = l, \hat X_t = j\,] \hat P_{jl}\\  
  & =
  \sum_l P [\,h_{t+T}(C_k) < h_{t+T}(\cup_{l\neq k} C_l)\vert \hat X_{t+T} = l\,] \hat P_{jl}\\
  &= 
  \sum_l  \hat P_{jl}\hat M_{lk}
\end{align*}
